---
title: "MERDs Data Analyses"
author: "Tayler Ulbrich and Lukas Bell-Dereske"
date: "9/27/2021"
output:
  word_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

# Load Files 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(ggpubr)
library(car)
library(emmeans)
library(dplyr)
library(MASS)
library(knitr)
library(here)
library(MuMIn)
library(lme4)
library(lmerTest)
options(contrasts=c("contr.sum", "contr.poly"))

germ.md <- read.csv(here::here("R_data","MERDS_GerminationExperiment_Rerun_Metadata.csv"), header = TRUE)

germ.raw <- read.csv(here::here("R_data","MERDS_GerminationExperiment_Rerun_RAWdata.csv"), header = TRUE)

germdata <- merge(germ.md, germ.raw, by = "ID")
dim(germdata)




```
# Checking possible confounding factors
## Does location in the growth chamber affect the germination
```{r}
head(germdata)

ggplot(germdata,aes(x=Shelf,y=Day12, color=Placement,group=interaction(Shelf,Placement)))+geom_boxplot()
ggplot(germdata,aes(x=Shelf,y=Day12))+geom_boxplot()
ggplot(germdata,aes(x=Placement,y=Day12, color=factor(Placement)))+geom_boxplot()



qqPlot(germdata$Day12)
hist(germdata$Day12)
lm.GermDay12_pos = lm(log(Day12+1) ~ Shelf+Placement,data = germdata)
qqPlot(stdres(lm.GermDay12_pos))
hist(stdres(lm.GermDay12_pos))
shapiro.test(stdres(lm.GermDay12_pos)) 
#plot(lm.GermDay12_pos)

# ANOVA 
Anova(lm.GermDay12_pos, type = "III")
#Shelf has a significant effect
```

# Day 12 Germination 
## PLOT
```{r}


germdata_Day12_sum=germdata %>% group_by(Soil_Treatment,Nitrogen_Treatment) %>% 
  summarise(mean=mean(Day12),n=n(),se=sd(Day12)/sqrt(n()))

(germination_Day12_p=ggplot(germdata)+
    geom_errorbar(data = germdata_Day12_sum,aes(group=factor(Nitrogen_Treatment,levels = c("Low","High")),
                                                  x=factor(Soil_Treatment, levels = c("Sterile","Live")),y=mean,ymin=mean-se,ymax=mean+se),
                width=.3,size=1, position = position_dodge(0.5), color="black")+
  geom_point(aes(x=factor(Soil_Treatment, levels = c("Sterile","Live")),y=Day12,
                 shape=factor(Nitrogen_Treatment,levels = c("Low","High")),
                 group=factor(Nitrogen_Treatment,levels = c("Low","High"))), size=3, position = position_dodge(0.5),alpha=0.5,fill="white")+
  geom_point(data = germdata_Day12_sum,aes(x=factor(Soil_Treatment, levels = c("Sterile","Live")),y=mean, 
                                               shape=factor(Nitrogen_Treatment,levels = c("Low","High")),
                                          group=factor(Nitrogen_Treatment,levels = c("Low","High"))), 
             size=7, position = position_dodge(0.5),stroke=1.5,fill="white")+
  scale_y_continuous(name ="Germination per Petri dish\n(# of germinants)")+
  scale_x_discrete(label=c("Sterile","Live"))+
  scale_shape_manual(values= c(25,15),name="Nitrogen")+
  theme_bw(base_size = 24)+theme(axis.text = element_text(size = 30),axis.title.x = element_blank(), axis.title.y = element_text(size = 36),
                   panel.grid.major = element_blank(), panel.grid.minor = element_blank(),legend.key.height = unit(1,"cm")))

#ggsave(here::here("R_fig","Nit_germination_Day12_dot_error.png"),germination_Day12_p,width =12,height = 8,device="png")
#ggsave(here::here("R_fig","Nit_germination_Day12_dot_error.svg"),germination_Day12_p,width =12,height = 8,device="svg")
```
## Stats
```{r}


# test for normal residuals and equal variances (ANOVA assumptions)


# check normal residuals
qqPlot(germdata$Day12)
hist(germdata$Day12)
lm.GermDay12 = lm(log(Day12+1) ~ Soil_Treatment*Nitrogen_Treatment, data = germdata)

# shapiro test - null hypothesis: residuals are normally distributed 
hist(stdres(lm.GermDay12))
qqPlot(stdres(lm.GermDay12))
shapiro.test(stdres(lm.GermDay12)) 
#plot(lm.GermDay12)

lm.GermDay12.1 = lmer(log(Day12+1) ~ Soil_Treatment*Nitrogen_Treatment + (1|Shelf)+(1|Placement), data = germdata)
hist(resid(lm.GermDay12.1))
qqPlot(resid(lm.GermDay12.1))
shapiro.test(resid(lm.GermDay12.1)) 
plot(lm.GermDay12.1)
lm.GermDay12.2 = lmer(log(Day12+1) ~ Soil_Treatment*Nitrogen_Treatment + (1|Shelf), data = germdata)
hist(resid(lm.GermDay12.2))
qqPlot(resid(lm.GermDay12.2))
shapiro.test(resid(lm.GermDay12.2)) 
plot(lm.GermDay12.2)
lm.GermDay12.3 = lmer(log(Day12+1) ~ Soil_Treatment*Nitrogen_Treatment + (1|Placement), data = germdata)
hist(resid(lm.GermDay12.3))
qqPlot(resid(lm.GermDay12.3))
shapiro.test(resid(lm.GermDay12.3)) 
plot(lm.GermDay12.3)
lm.GermDay12.4 = lm(log(Day12+1) ~ Soil_Treatment*Nitrogen_Treatment + Shelf, data = germdata)
hist(resid(lm.GermDay12.4))
qqPlot(resid(lm.GermDay12.4))
shapiro.test(resid(lm.GermDay12.4)) 
#plot(lm.GermDay12.4)
lm.GermDay12.5 = lm(log(Day12+1) ~ Soil_Treatment*Nitrogen_Treatment + Shelf + Placement, data = germdata)
hist(stdres(lm.GermDay12.5))
qqPlot(stdres(lm.GermDay12.5))
shapiro.test(stdres(lm.GermDay12.5)) 
#plot(lm.GermDay12.5)
lm.GermDay12.6 = lm(log(Day12+1) ~ Soil_Treatment*Nitrogen_Treatment + Placement, data = germdata)
hist(stdres(lm.GermDay12.6))
qqPlot(stdres(lm.GermDay12.6))
shapiro.test(stdres(lm.GermDay12.6)) 
#plot(lm.GermDay12.6)

AICc(lm.GermDay12,lm.GermDay12.1,lm.GermDay12.2,lm.GermDay12.3,lm.GermDay12.4,lm.GermDay12.5,lm.GermDay12.6)



# ANOVA 
Anova(lm.GermDay12.4, type = "III")
#Soil_Treatment                     0.7233  1  3.5921   0.06634 .  
#Shelf                              1.3805  1  6.8561   0.01297 *  


```
Formatted Anova table
```{r echo=FALSE, results='asis'}
anova_lm.GermDay12=Anova(lm.GermDay12.4, type = "III")
df_anova_lm.GermDay12=data.frame(row.names =row.names(anova_lm.GermDay12),"Df"=anova_lm.GermDay12$Df,"F-value"=anova_lm.GermDay12$`F value`,"P-value"=anova_lm.GermDay12$`Pr(>F)`)
kable(df_anova_lm.GermDay12,align = "c",digits = c(1,2,3), caption= "Germination at day 3")
```

#### Percentage Difference and magnitude High versus low nitrogen
```{r echo=FALSE}
#Mean germination across treatments
germdata %>%  summarise_at("Day12", list(~n(),~mean(.),~sd(.),se=~sd(.)/sqrt(n())))

germdata_N_level=germdata %>% group_by(Nitrogen_Treatment)%>% summarise_at("Day12", list(~n(),~mean(.),~sd(.),se=~sd(.)/sqrt(n())))

#Percentage Difference
((germdata_N_level[germdata_N_level$Nitrogen_Treatment=="High",]$mean-germdata_N_level[germdata_N_level$Nitrogen_Treatment=="Low",]$mean)/germdata_N_level[germdata_N_level$Nitrogen_Treatment=="Low",]$mean)*100


germdata_N_level[germdata_N_level$Nitrogen_Treatment=="High",]$mean-germdata_N_level[germdata_N_level$Nitrogen_Treatment=="Low",]$mean

```






## Does location in the growth chamber affect the per germinant biomass?
```{r}
head(germdata)

summary(germdata)

#Convert NA's to 0's
germdata$TotalGerminant_shoot_dry_g=ifelse(is.na(germdata$TotalGerminant_shoot_dry_g),
                                           0,germdata$TotalGerminant_shoot_dry_g)
germdata$PerGerminant_shoot_dry_g=ifelse(germdata$TotalGerminant_shoot_dry_g==0,0,
                                         germdata$TotalGerminant_shoot_dry_g/germdata$Day12)



ggplot(germdata,aes(x=Shelf,y=PerGerminant_shoot_dry_g, color=Placement,group=interaction(Shelf,Placement)))+geom_boxplot()
ggplot(germdata,aes(x=Shelf,y=PerGerminant_shoot_dry_g))+geom_boxplot()
ggplot(germdata,aes(x=Placement,y=PerGerminant_shoot_dry_g, color=factor(Placement)))+geom_boxplot()



qqPlot(germdata$PerGerminant_shoot_dry_g)
hist(germdata$PerGerminant_shoot_dry_g)
lm.TotalBio_pos = lm(sqrt(PerGerminant_shoot_dry_g+0.0001) ~ Shelf+Placement,data = germdata)
qqPlot(stdres(lm.TotalBio_pos))
hist(stdres(lm.TotalBio_pos))
shapiro.test(stdres(lm.TotalBio_pos)) 
#plot(lm.GermDay12_pos)

# ANOVA 
Anova(lm.TotalBio_pos, type = "III")
#Niether has a significnat effect
```

# Day 12 Biomass 
## PLOT
```{r}


germdata_PerGerminant_shoot_dry_g_sum=germdata %>% group_by(Soil_Treatment,Nitrogen_Treatment) %>% 
  summarise(mean=mean(PerGerminant_shoot_dry_g),n=n(),se=sd(PerGerminant_shoot_dry_g)/sqrt(n()))

(germination_PerGerminant_shoot_dry_g_p=ggplot(germdata)+
    geom_errorbar(data = germdata_PerGerminant_shoot_dry_g_sum,aes(group=factor(Nitrogen_Treatment,levels = c("Low","High")),
                                                  x=factor(Soil_Treatment, levels = c("Sterile","Live")),y=mean,ymin=mean-se,ymax=mean+se),
                width=.3,size=1, position = position_dodge(0.5), color="black")+
  geom_point(aes(x=factor(Soil_Treatment, levels = c("Sterile","Live")),y=PerGerminant_shoot_dry_g,
                 shape=factor(Nitrogen_Treatment,levels = c("Low","High")),
                 group=factor(Nitrogen_Treatment,levels = c("Low","High"))), size=3, position = position_dodge(0.5),alpha=0.5,fill="white")+
  geom_point(data = germdata_PerGerminant_shoot_dry_g_sum,aes(x=factor(Soil_Treatment, levels = c("Sterile","Live")),y=mean, 
                                               shape=factor(Nitrogen_Treatment,levels = c("Low","High")),
                                          group=factor(Nitrogen_Treatment,levels = c("Low","High"))), 
             size=7, position = position_dodge(0.5),stroke=1.5,fill="white")+
  scale_y_continuous(name ="Biomass per germinant (g)")+
  scale_x_discrete(label=c("Sterile","Live"))+
  scale_shape_manual(values= c(25,15),name="Nitrogen")+
  theme_bw(base_size = 24)+theme(axis.text = element_text(size = 30),axis.title.x = element_blank(), axis.title.y = element_text(size = 36),
                   panel.grid.major = element_blank(), panel.grid.minor = element_blank(),legend.key.height = unit(1,"cm")))

#ggsave(here::here("R_fig","Nit_germination_PerGerminant_shoot_dry_dot_error.png"),germination_PerGerminant_shoot_dry_g_p,width =12,height = 8,device="png")
#ggsave(here::here("R_fig","Nit_germination_PerGerminant_shoot_dry_dot_error.svg"),germination_PerGerminant_shoot_dry_g_p,width =12,height = 8)
```
## Stats
```{r}


# test for normal residuals and equal variances (ANOVA assumptions)


# check normal residuals
qqPlot(germdata$PerGerminant_shoot_dry_g)
hist(germdata$PerGerminant_shoot_dry_g)
lm.GermPerGerminant_shoot_dry_g = lm(sqrt(PerGerminant_shoot_dry_g+0.0001) ~ Soil_Treatment*Nitrogen_Treatment, data = germdata)

# shapiro test - null hypothesis: residuals are normally distributed 
hist(stdres(lm.GermPerGerminant_shoot_dry_g))
qqPlot(stdres(lm.GermPerGerminant_shoot_dry_g))
shapiro.test(stdres(lm.GermPerGerminant_shoot_dry_g)) 
#plot(lm.GermPerGerminant_shoot_dry_g)

lm.GermPerGerminant_shoot_dry_g.1 = lmer(sqrt(PerGerminant_shoot_dry_g+0.0001) ~ Soil_Treatment*Nitrogen_Treatment + (1|Shelf)+(1|Placement), data = germdata)
hist(resid(lm.GermPerGerminant_shoot_dry_g.1))
qqPlot(resid(lm.GermPerGerminant_shoot_dry_g.1))
shapiro.test(resid(lm.GermPerGerminant_shoot_dry_g.1)) 
plot(lm.GermPerGerminant_shoot_dry_g.1)
lm.GermPerGerminant_shoot_dry_g.2 = lmer(sqrt(PerGerminant_shoot_dry_g+0.0001) ~ Soil_Treatment*Nitrogen_Treatment + (1|Shelf), data = germdata)
hist(resid(lm.GermPerGerminant_shoot_dry_g.2))
qqPlot(resid(lm.GermPerGerminant_shoot_dry_g.2))
shapiro.test(resid(lm.GermPerGerminant_shoot_dry_g.2)) 
plot(lm.GermPerGerminant_shoot_dry_g.2)
lm.GermPerGerminant_shoot_dry_g.3 = lmer(sqrt(PerGerminant_shoot_dry_g+0.0001) ~ Soil_Treatment*Nitrogen_Treatment + (1|Placement), data = germdata)
hist(resid(lm.GermPerGerminant_shoot_dry_g.3))
qqPlot(resid(lm.GermPerGerminant_shoot_dry_g.3))
shapiro.test(resid(lm.GermPerGerminant_shoot_dry_g.3)) 
plot(lm.GermPerGerminant_shoot_dry_g.3)
lm.GermPerGerminant_shoot_dry_g.4 = lm(sqrt(PerGerminant_shoot_dry_g+0.0001) ~ Soil_Treatment*Nitrogen_Treatment + Shelf, data = germdata)
hist(resid(lm.GermPerGerminant_shoot_dry_g.4))
qqPlot(resid(lm.GermPerGerminant_shoot_dry_g.4))
shapiro.test(resid(lm.GermPerGerminant_shoot_dry_g.4)) 
#plot(lm.GermPerGerminant_shoot_dry_g.4)
lm.GermPerGerminant_shoot_dry_g.5 = lm(sqrt(PerGerminant_shoot_dry_g+0.0001) ~ Soil_Treatment*Nitrogen_Treatment + Shelf + Placement, data = germdata)
hist(stdres(lm.GermPerGerminant_shoot_dry_g.5))
qqPlot(stdres(lm.GermPerGerminant_shoot_dry_g.5))
shapiro.test(stdres(lm.GermPerGerminant_shoot_dry_g.5)) 
#plot(lm.GermPerGerminant_shoot_dry_g.5)
lm.GermPerGerminant_shoot_dry_g.6 = lm(sqrt(PerGerminant_shoot_dry_g+0.0001) ~ Soil_Treatment*Nitrogen_Treatment + Placement, data = germdata)
hist(stdres(lm.GermPerGerminant_shoot_dry_g.6))
qqPlot(stdres(lm.GermPerGerminant_shoot_dry_g.6))
shapiro.test(stdres(lm.GermPerGerminant_shoot_dry_g.6)) 
#plot(lm.GermPerGerminant_shoot_dry_g.6)

AICc(lm.GermPerGerminant_shoot_dry_g,lm.GermPerGerminant_shoot_dry_g.1,lm.GermPerGerminant_shoot_dry_g.2,lm.GermPerGerminant_shoot_dry_g.3,lm.GermPerGerminant_shoot_dry_g.4,lm.GermPerGerminant_shoot_dry_g.5,lm.GermPerGerminant_shoot_dry_g.6)



# ANOVA 
Anova(lm.GermPerGerminant_shoot_dry_g.4, type = "III")
#Soil_Treatment                    0.0005444  1   7.9478  0.007872 ** 
#Shelf                             0.0002237  1   3.2656  0.079350 . 


```
Formatted Anova table
```{r echo=FALSE, results='asis'}
anova_lm.GermPerGerminant_shoot_dry_g=Anova(lm.GermPerGerminant_shoot_dry_g.4, type = "III")
df_anova_lm.GermPerGerminant_shoot_dry_g=data.frame(row.names =row.names(anova_lm.GermPerGerminant_shoot_dry_g),"Df"=anova_lm.GermPerGerminant_shoot_dry_g$Df,"F-value"=anova_lm.GermPerGerminant_shoot_dry_g$`F value`,"P-value"=anova_lm.GermPerGerminant_shoot_dry_g$`Pr(>F)`)
kable(df_anova_lm.GermPerGerminant_shoot_dry_g,align = "c",digits = c(1,2,3), caption= "Germination at day 3")
```

#### Percentage Difference and magnitude High versus low nitrogen
```{r echo=FALSE}
#Mean germination across treatments
germdata %>%  summarise_at("PerGerminant_shoot_dry_g", list(~n(),~mean(.),~sd(.),se=~sd(.)/sqrt(n())))

germdata_N_level_biomass=germdata %>% group_by(Nitrogen_Treatment)%>% summarise_at("PerGerminant_shoot_dry_g", list(~n(),~mean(.),~sd(.),se=~sd(.)/sqrt(n())))

#Percentage Difference
((germdata_N_level_biomass[germdata_N_level_biomass$Nitrogen_Treatment=="High",]$mean-germdata_N_level_biomass[germdata_N_level_biomass$Nitrogen_Treatment=="Low",]$mean)/germdata_N_level_biomass[germdata_N_level_biomass$Nitrogen_Treatment=="Low",]$mean)*100


germdata_N_level_biomass[germdata_N_level_biomass$Nitrogen_Treatment=="High",]$mean-germdata_N_level_biomass[germdata_N_level_biomass$Nitrogen_Treatment=="Low",]$mean


germdata_soil_status_biomass=germdata %>% group_by(Soil_Treatment)%>% summarise_at("PerGerminant_shoot_dry_g", list(~n(),~mean(.),~sd(.),se=~sd(.)/sqrt(n())))

#Percentage Difference
((germdata_soil_status_biomass[germdata_soil_status_biomass$Soil_Treatment=="Sterile",]$mean-germdata_soil_status_biomass[germdata_soil_status_biomass$Soil_Treatment=="Live",]$mean)/germdata_soil_status_biomass[germdata_soil_status_biomass$Soil_Treatment=="Live",]$mean)*100


germdata_soil_status_biomass[germdata_soil_status_biomass$Nitrogen_Treatment=="High",]$mean-germdata_soil_status_biomass[germdata_soil_status_biomass$Nitrogen_Treatment=="Low",]$mean

```













